The association between Facebook adoption and well-being

### 2022-09-16 data-prepare.Rmd
rename

## Data cleaning

@@

@@

### Population

Yearly population estimates from world bank, except Taiwan from Wikipedia.

```{r}
pop <- read_csv(
  "data/API_SP.POP.TOTL_DS2_en_csv_v2_4335082.csv", 
  skip = 3
)
pop <- pop %>% 
  select(
    country = `Country Code`,
    `2008`:`2019`
  ) %>% 
  mutate(
    country = countrycode(country, origin = "iso3c", destination = "cldr.short.en")
    ) %>% 
  filter(country %in% fb_countries) %>% 
  pivot_longer(`2008`:`2019`, names_to = "year", values_to = "pop") %>% 
  mutate(year = as.numeric(year))

tw <- tibble(
  country = "Taiwan",
  year = 2008:2019,
  pop = c(23037031, 23119772, 23162123, 23224912, 23315822, 23373517, 23433753, 23492074, 23539816, 23571227, 23588932, 23600903)
)
pop <- bind_rows(pop, tw)
```

>>>>

# Data cleaning

@@

## Gallup

@@

## Population

From <https://population.un.org/wpp/Download/Standard/Population/>.

```{r}
library(readxl)
pop <- read_excel(
  "data/WPP2022_POP_F01_1_POPULATION_SINGLE_AGE_BOTH_SEXES.xlsx", 
  sheet = 1, 
  skip = 16,
  col_types = "text",
  .name_repair = make_clean_names
  )
pop <- pop %>% 
  drop_na(iso3_alpha_code) %>% 
  select(
    iso3_alpha_code, year, starts_with("x")
  ) %>% 
  mutate(
    country = countrycode(
      iso3_alpha_code, 
      origin = "iso3c", 
      destination = "cldr.short.en"
    )
  )
pop <- pop %>% 
  filter(
    country %in% fb_countries,
    between(year, 2008, 2019)
  ) %>% 
  pivot_longer(
    starts_with("x"), 
    names_to = "age",
    values_to = "pop",
    names_transform = ~str_remove(., "x") %>% as.numeric, 
    values_transform = ~ as.numeric(.) * 1000
  ) %>% 
  filter(between(age, 13, 100)) %>% 
  mutate(
    age = cut(age, c(0, 34.5, 1000), labels = c("13-34", "35+"))
  )
pop <- pop %>% 
  mutate(year = as.numeric(year)) %>% 
  group_by(country, year, age) %>% 
  summarise(pop = sum(pop)) %>% 
  ungroup()
```

## Synthetic Facebook data

Here we create simple time trend data that can be used for reproducing the computations when the actual Facebook data is not available. It does not represent any statistical properties of the Facebook data and is created simply to help run the computations.

```{r}
fb_mock_path <- here("data/fb-SYNTHETIC.rds")
if (!file.exists(fb_mock_path)) {
  library(extraDistr)
  set.seed(999)
  slopes <- distinct(pop, country) %>% 
    mutate(slope = rnorm(72, 0.75, 0.1))
  fb_syn <- pop %>% 
    left_join(slopes) %>% 
    rowwise() %>% 
    mutate(
      dau = rprop(1, 50, plogis(-4 + slope*(year-2008))),
      mau = rprop(1, 100, plogis(-5 + slope*(year-2008)))
    )
  saveRDS(fb_syn, fb_mock_path)
}
```
